---
title: "Replication_Huh2016_PilotB"
author: "Danielle Boles"
date: "12/6/2017"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse)
library(DT)
library(tidyr)
library(ggplot2)
library(knitr)
library(broom)
library(devtools)
library(lme4)
library(ggplot2)

```

<center> Replication of "More Similar but Less Satisfying: Comparing Preferences for and the Efficacy of Within- and Cross-Category Substitutes for Food" by Huh, Vosgerau, and Morewedge (2016, *Psychological Science*) </center>

<center>Danielle Boles <br>
Department of Psychology <br>
Stanford University <br>
dboles@stanford.edu</center>

\newpage

<center>
##Introduction
</center>

The original article by Huh, Vosgerau and Morewedge examines choice behavior and expectations among within- and cross-category substitutes when their desired food is unavailable. This report seeks to replicate Study 1 findings, which found that people more often chose a within-category substitute (store brand chocolate) over a cross-category substitute (granola bar) when their desired food (gourmet chocolate) was not available, despite their expectation that within-category substitues will be less enjoyable. This finding suggests that restricting the availability of desired foods negatively influences our expectations of similar foods due to a negative contrast effect.

## Methods
> [Link to survey](https://stanforduniversity.qualtrics.com/jfe/form/SV_7Pe4dqmYThp3OrH) 

### Power Analysis
The key statistical analyses by which the success or failure of this replication will be judged is the significance of the finding that even though participants expected to enjoy within-category (M = 2.86, SD = 0.90) substitutes less than cross-category substitutes (M = 3.48, SD = 0.81), all t(100)s > 7.53, all ps < .001, participants more often chose within-category substitutes than cross-category substitutes (mean choice share for within-category substitutes = 79.7%)...test against 50%: p < .001.
The original article sampled 101 online participants, which when performing a posthoc power calculation on enjoyment ratings, yielded an effect size of d = .45 with 99% power. In terms of numbers of partcipants, 90% power will require N=44, less than half of the original sample. For the purposes of this replication, it will seek to sample 55 participants, achieving 95% power.

### Planned Sample
Online participants will be recruited via Amazon's Mechanical Turk and must be at least 18 years old, have an approval rating of 95%, and live in the United States. Demographic variables of the original study were not reported. Although, sample size calculations show that we would only need 32 participants to meet the threshold of 80% power, this replication will sample 100 participants for two reasons:
1. To match the high power of the original study.
2. The estimated time of survey completion is only 3-4 min.
Apart from the age, approval rating, and consent requirements, the authors did not list any exclusionary criteria in creating their final sample.

### Materials
This survey was created on Qualtrics and launched on Amazon Mechanical Turk. Participants are required to complete this survey on a laptop or tablet.

### Procedure
> "Each participant saw 10 separate sets of three foods: one desired food and two substiutes. For each set, they first indicated which of two foods would vetter satisfy their craving for the desired food if it were not available: a within-category substitute or a cross-category substitute (e.g., if the desired food was a "Chipotle burrito," they could choose a "store-brand frozen burrito" or "KFC fried chicken" as the within- and cross-category substitute, respectively... Sets and choice options within sets were presented in random order. Stimuli were chosen such that within-category substitutes were more similar than cross category substitutes to the desired foods but were expected to be less enjoyable to consume (i.e., of lower quality). Participants were given the following information before seeing the stimulus sets:
In some cases, we describe foods as a “store-brand” or “gas-station-brand” food. Please assume these foods are store-brand foods sold by your local supermarket, drug store, or gas station. For example, CVS milk chocolate would be a store-brand food.
Participants then indicated how much they expected they would enjoy eating each food in the set (i.e., desired food, within-category substitute, and cross-category substitute) and rated which of the two substitutes was more similar to the desired food. Participants rated the extent to which they expected to enjoy each food on a 5-point scale from 1, not at all, to 5, very much. Participants rated which of the two substitutes was more similar to the desired food on a 7-point scale with values closer to 1 indicating that the within-category substitute was more similar to the desired food, values closer to 7 indicating that the cross- category substitute was more similar to the desired food, and values at the midpoint (4) indicating that the two sub- stitutes were equally similar to the desired food. Rating order was counterbalanced across participants."



### Analysis Plan
This replication will first perform a manipulation check by testing each similarity rating against the scale midpoint (4).
with a paired samples t-test. 
For the key analysis of interest, the overall enjoyment ratings will also be calculated with t-tests to show differences in enjoyment among cross category, within category, and target items.
Mean choice share was calculated for each substitute and tested against 50%. Choices will be regressed on similarity and enjoyment ratings with participants as random effects using Generalized linear mixed model fit by maximum likelihood (glmer). The authors were unclear which tests they ran for each statistic, so the analysis of this replication was determined by doing the following:
1. Trying to reproduce the findings using SPSS.
2. 'Translating' SPSS analyses parameters into R.
3. Writing the R script & testing it on original data to reproduce findings.
Where this replication differs in analysis reporting (i.e., z scores, Wald chi square) is noted below.


#### Importing the Data
```{r}
setwd("~/Desktop/PSYCH 251 - Stats/Huh2016/data")
d <- read.csv("Replication_Huh2016_PilotA.csv")

```

#### Cleaning the Data
```{r}
# cleaning the data, reclassifying variable types
d <- d %>%
        select((starts_with("sub")), (starts_with("pref")), (starts_with("sim")), "Gender", "Age") %>%
        mutate_at(vars(contains("sub")), funs(as.factor)) %>%
        mutate_at(vars(contains("sim")), funs(as.numeric)) %>%
        mutate_at(vars(contains("pref")), funs(as.numeric))

d$sub1 <- factor(d$sub1, labels = c("cross category", "within category"))
d$sub2 <- factor(d$sub2, labels = c("cross category", "within category"))
d$sub3 <- factor(d$sub3, labels = c("cross category", "within category"))
d$sub4 <- factor(d$sub4, labels = c("cross category", "within category"))
d$sub5 <- factor(d$sub5, labels = c("cross category", "within category"))
d$sub6 <- factor(d$sub6, labels = c("cross category", "within category"))
d$sub7 <- factor(d$sub7, labels = c("cross category", "within category"))
d$sub8 <- factor(d$sub8, labels = c("cross category", "within category"))
d$sub9 <- factor(d$sub9, labels = c("cross category", "within category"))
d$sub10 <- factor(d$sub10, labels = c("cross category", "within category"))

# add subject id column
id <- rownames(d)
d <- cbind(id=id, d)


head(d)

```

#### Manipulation Check (similarity rating) 
```{r}

# t tests for each set of stimuli
t.test(d$sim1, mu=4)
t.test(d$sim2, mu=4)
t.test(d$sim3, mu=4)
t.test(d$sim4, mu=4)
t.test(d$sim5, mu=4)
t.test(d$sim6, mu=4)
t.test(d$sim7, mu=4)
t.test(d$sim8, mu=4)
t.test(d$sim9, mu=4)
t.test(d$sim10, mu=4)

# convert to tidy data
d_tidy_sim <- d %>%
        select(id, starts_with("sim")) %>%
        gather(set, sim, starts_with("sim"))

# cleaning the data
d_tidy_sim$set <- gsub("sim", "", d_tidy_sim$set)
d_tidy_sim$set <- as.numeric(d_tidy_sim$set)
d_tidy_sim$set <- as.factor(d_tidy_sim$set)

head(d_tidy_sim)

#calculating mean and sd

d_tidy_sim_mean <- d_tidy_sim %>%
        group_by(set) %>%
        summarise(mean = mean(sim, na.rm = TRUE))

d_tidy_sim_sd <- d_tidy_sim %>%
        group_by(set) %>%
        summarise(sd = sd(sim, na.rm = TRUE))

# creating table of mean & sd
d_tidy_sim_mean_sd <- full_join(d_tidy_sim_mean, d_tidy_sim_sd, by = NULL, copy = FALSE)

head(d_tidy_sim_mean_sd)

kable(d_tidy_sim_mean_sd, digits = 2)

# calculating grand mean
sim_grandmean = (mean(d_tidy_sim_mean_sd$mean))
sim_grandsd = (mean(d_tidy_sim_mean_sd$sd))

sim_grandmean
sim_grandsd

```
Each of the 10 similarity ratings was tested against the scale midpoint (4; they are equally similar). Participants judged all within-category substitutes to be more similar than all cross-category substitutes to the desired foods (M = 2.38, SD = 1.34, Range = 2.1 - 2.9), all ps < .05 except for stimulus sets #3 (p <.1) and #5 (p < .1).


#### Key Statistical Analysis 1: Expected Enjoyment (Preference) Ratings
```{r}

# Calculating mean enjoyment and sd of substitute item type by stimulus set
d_tidy_pref <- d %>%
        select(id, starts_with("pref")) %>%
        gather(pref, pref_rating, starts_with("pref")) %>%
        separate(pref, c("set", "item"), "_") %>%
        spread(item, pref_rating) %>%
        rename(within = "1",
               cross = "2",
               target = "3") %>%
        select(set, within, cross, target) %>%
        group_by(set) %>%
        summarise(
                mean_within = mean(within, na.rm = TRUE, round(2)),
                mean_cross = mean(cross, na.rm = TRUE, round(2)),
                mean_target = mean(target, na.rm = TRUE, round(2))) %>%
        mutate(mean_within, mean_cross, mean_target)

# cleaning the data
d_tidy_pref$set <- gsub("pref", "", d_tidy_pref$set)
d_tidy_pref$set <- as.factor(d_tidy_pref$set)

head(d_tidy_pref)
kable(d_tidy_pref)

# **NEED HELP calculating 3 t-tests for each set. 1 t test each for within vs. cross, within vs. target, cross vs. target**

# Calculating t-tests by substitute item type overall
t.test(d_tidy_pref$mean_cross, d_tidy_pref$mean_within)
t.test(d_tidy_pref$mean_target, d_tidy_pref$mean_within)
t.test(d_tidy_pref$mean_cross, d_tidy_pref$mean_target)


# Creating table for overall mean enjoyment and sd of substitute item type
d_tidy_pref_overall <- d %>%
        select(id, starts_with("pref")) %>%
        gather(pref, pref_rating, starts_with("pref")) %>%
        separate(pref, c("set", "item"), "_") %>%
        group_by(item) %>%
        summarise(
                mean = mean(pref_rating, na.rm = TRUE),
                sd = sd(pref_rating, na.rm = TRUE)) %>%
        mutate(mean, sd)

d_tidy_pref_overall$item <- factor(c("within", "cross", "target"))

kable(d_tidy_pref_overall, digits = 2)

# Graphing means & sd by item
ggplot(d_tidy_pref_overall, aes(x=item, y=mean, fill=item)) +
  geom_bar(stat="identity") +
        geom_errorbar(
                data = d_tidy_pref_overall,
                aes(item, mean, ymin = mean - sd, ymax = mean + sd),
                colour = "black",
                width = 0.4) +
        labs(title="Mean Enjoyment Ratings by Substitute Item",
       x="Substitute Item", y = "Mean Enjoyment")
```

On this pilot data, participants overall expected to enjoy eating the target item (M = 3.99, SD = 1.18) and the cross category item (M = 3.80, SD = 1.37) more than the within category item (M = 2.20, SD = 0.99), p>.001. There was no significant difference between the enjoyment ratings for the target item and the enjoyment ratings for the cross-category item (p>.05). This replication diverges from the original study by reporting a t-test on all enjoyment ratings of items (target vs. within, target vs. cross, within vs. cross) across all sets of stimuli (Set 1, Set 2,... Set 10) whereas the original authors did t-tests for each stimuli set and reported all t(100s) > 7.53 and all ps <.001.
        

#### Key Statistical Analysis 2: Item Choice Share
```{r}

# gathering data. 0 = cross category, 1 = within category
d_tidy_choice <- d %>%
        select(id, starts_with("sub")) %>%
        gather(set, choice, starts_with("sub"))

# cleaning the data
d_tidy_choice$set <- gsub("sub", "", d_tidy_choice$set)
d_tidy_choice$choice <- gsub("cross category", "1", d_tidy_choice$choice)
d_tidy_choice$choice <- gsub("within category", "2", d_tidy_choice$choice)

head(d_tidy_choice)

# calculate proportions chose within-category vs. cross-category & test against 50%
table(d_tidy_choice$choice)
binom.test(41, 50, 0.5, alternative="two.sided")
```
Overall, compared to within category substitutes, participants more often chose cross category items to satisfy their craving for the target food. 
This replication diverges from the original paper by not reporting a z-score. The original authors reported a z score of 5.97. 




